---
title: "Simulate data to test shiny app"
format: 
  html:
    toc: true
    toc-location: left
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12,
                      fig.height=8,
                      results='verbatim',
                      cache=TRUE,
                      message=F,
                      warning=F,
                      progress=FALSE,
                      echo=TRUE)
```

```{r}
library(mvtnorm)
library(Matrix)
library(tidyverse)
library(magrittr)
library(bayesforecast)
library(rstanarm)
library(ggplot2)
library(plotly)
library(factoextra)
library(flextable)
library(bridgesampling)

sim_data = list()
fits = list()
```


# Cross-sectional no regression

## Simulate data
Note to SB: The covariance matrices derived from the outer(...) just correspond to an AR(1) process with different correlation parameters.   
```{r}
sim_data$crosssec_no_regr = 
  rmvnorm(100,
          mean = numeric(5),
          sigma = bdiag(outer(1:2,1:2,
                              function(x,y){0.6^abs(x-y)}),
                        outer(1:3,1:3,
                              function(x,y){(-0.2)^abs(x-y)})) %>% 
            as.matrix()) %>% 
  rbind(rmvnorm(200,
                mean = numeric(5),
                sigma = bdiag(outer(1:3,1:3,
                                    function(x,y){0.8^abs(x-y)}),
                              outer(1:2,1:2,
                                    function(x,y){(0.1)^abs(x-y)})) %>% 
                  as.matrix())) %>% 
  as_tibble() %>% 
  rename_all(function(x) gsub("V","y",x)) %>% 
  mutate(grouping_variable = rep(c("G1","G2"),c(100,200)))
```

## Univariate analyses
```{r}
cont_vars = 
  colnames(sim_data$crosssec_no_regr)[which(sim_data$crosssec_no_regr %>% 
                                              sapply(class) %in%
                                              c("numeric","integer"))]
cat_vars = 
  colnames(sim_data$crosssec_no_regr)[which(sim_data$crosssec_no_regr %>% 
                                              sapply(class) %in%
                                              c("character","factor"))]

plots = list(cont = list(),
             cat = list())

for(v in cont_vars){
  plots$cont[[v]] = 
    sim_data$crosssec_no_regr %>% 
    ggplot(aes_string(x = v)) +
    geom_density(adjust = 2,
                 fill = "steelblue") +
    labs(x = v, 
         y = "") +
    theme_classic() +
    theme(axis.title.x = element_text(size = 20),
          axis.title.y = element_text(size = 15),
          axis.text.x = element_text(size = 18),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank())
}
for(v in cat_vars){
  plots$cat[[v]] = 
    sim_data$crosssec_no_regr %>% 
      ggplot(aes_string(x = v)) +
      geom_bar(fill = "steelblue") +
      labs(x = "Grouping variable", 
           y = "") +
      theme_classic() +
      theme(axis.title.x = element_text(size = 20),
            axis.title.y = element_text(size = 15),
            axis.text.x = element_text(size = 18),
            axis.text.y = element_text(size = 18))
}

for(i in names(plots)){
  for(j in names(plots[[i]]))
  print(plots[[i]][[j]])
}

```

## Multivariate
Principle Components Analysis:
```{r}
data_pca = 
  sim_data$crosssec_no_regr %>% 
  select(all_of(cont_vars)) %>% 
  princomp()
```

### Screeplot
```{r}
data_pca %>% 
  fviz_screeplot() +
  theme_classic() +  # Use a classic theme
  theme(
    plot.title = element_text(size = 20),  # Increase title size
    axis.title = element_text(size = 20),  # Increase axis labels size
    axis.text = element_text(size = 18)    # Increase axis text size
  )
```

### Variables only
You can play around with the arguments here to vary the appearance, increase axis labels, etc.
```{r}
data_pca %>% 
  fviz_pca_var(col.var="contrib",
               gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
               repel = TRUE )
```

### Observations only
```{r}
data_pca %>% 
  fviz_pca_ind(habillage = sim_data$crosssec_no_regr$grouping_variable,
               label = "none")
```



# Retro cohort no regression

## Simulate data
```{r}
sim_data$retro_no_regr = 
  tibble(subject_id = 
           rep(1:100, each = 5),
         time = rep(1:5,100)) %>% 
  group_by(subject_id) %>% 
  mutate(y1 = rbinom(n(),1,
                     1.0 / (1.0 + exp(2 - 0.5 * time)) ),
         y2 = rnorm(n(),
                    2 * (-1)^(time > 2))) %>% 
  ungroup()
```

## Spaghetti plots
```{r}
sim_data$retro_no_regr %>% 
  ggplot(aes(x = time,
             y = y1,
             group = subject_id)) +
  geom_jitter(height = 0.05,width = 0.05) +
  geom_line() +
  theme_classic() +
  theme(axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18))
```
```{r}
sim_data$retro_no_regr %>% 
  ggplot(aes(x = time,
             y = y2,
             group = subject_id)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  theme(axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18))

```




# Prospective cohort no regression
Same as above.



# Case-control (Unmatched)

## Simulate data
```{r}
sim_data$case_control = 
  tibble(status = 
           rep(c("control","case"),
               c(500, 100))) %>%
  mutate(covariate = rnorm(n()),
         outcome = 
           rbinom(600,
                  1,
                  rep(c(0.2,0.4),
                      c(500,100))))
```

## Analysis - Logistic Regression
```{r results = "hide"}
set.seed(2024)
fits$case_control = 
  stan_glm(outcome ~ covariate + status,
           data = sim_data$case_control,
           family = "binomial")
```

```{r}
tibble(Covariate = 
         names(coef(fits$case_control)),
       Estimate = coef(fits$case_control)) %>% 
  bind_cols(as_tibble(posterior_interval(fits$case_control,
                                         prob = 0.95))) %>% 
  flextable() %>% 
  colformat_double(digits = 3) %>% 
  theme_booktabs()
```



# Case-control (Matched)

## Simulate data
```{r}
logit = function(x) log(x / (1.0 - x))
expit = function(x) 1.0 / (1.0 + exp(-x))

sim_data$case_control_matched = 
  tibble(status = 
           rep(rep(c("control","case"),
                   c(5,1)),
               100),
         group_id = rep(1:100,each = 5+1),
         latent_random_effect = rep(rnorm(100,sd=0.05),each = 5+1)) %>%
  mutate(covariate = rnorm(n()),
         outcome = 
           rbinom(600,
                  1,
                  expit( logit(0.2) + # Intercept
                           (logit(0.4) - logit(0.2)) * (status == "case") +# Slope
                           latent_random_effect) ) )
```

## Analysis - Conditional Logistic Regression
```{r results = "hide"}
set.seed(2024)
fits$case_control_matched = 
  stan_clogit(outcome ~ covariate + status,
              strata = group_id,
              data = sim_data$case_control_matched)
```


```{r}
tibble(Covariate = 
         names(coef(fits$case_control_matched)),
       Estimate = coef(fits$case_control_matched)) %>% 
  bind_cols(as_tibble(posterior_interval(fits$case_control_matched,
                                         prob = 0.95))) %>% 
  flextable() %>% 
  colformat_double(digits = 3) %>% 
  theme_booktabs()
```



# Cross-sectional regression

## Continuous
### Simulate data
```{r}
sim_data$crosssec_regr = list()
sim_data$crosssec_regr$continuous = 
  tibble(x1 = rnorm(100),
         x2 = rnorm(100)) %>% 
  mutate(log_y = -1 + 1.5*x1 + rnorm(n())) %>% 
  mutate(y = exp(log_y))
```

### Analysis - linear regression
```{r results = "hide"}
set.seed(2024)
fits$linreg = list()
fits$linreg$untransformed = 
  stan_glm(y ~ x1 + x2,
           data = sim_data$crosssec_regr$continuous,
           diagnostic_file = file.path(tempdir(), "delete_me-1.csv"))

transformation = NULL
# Try transformations.
if(all(sim_data$crosssec_regr$continuous$y > 0)){
  set.seed(2024)
  fits$linreg$log = 
    stan_glm(log_y ~ x1 + x2,
             data = 
               sim_data$crosssec_regr$continuous %>% 
               mutate(log_y = log(y)),
             diagnostic_file = file.path(tempdir(), "delete_me-2.csv"))
  transformation = "log"
}else{
  if(all(sim_data$crosssec_regr$continuous$y >= 0)){
  set.seed(2024)
  fits$linreg$asinh = 
    stan_glm(asinh_y ~ x1 + x2,
             data = 
               sim_data$crosssec_regr$continuous %>% 
               mutate(asinh_y = asinh(y)),
           )
    transformation = "asinh"
  }
}

if(length(fits$linreg) > 1){
  set.seed(2024)
  bf = 
    bayes_factor(bridge_sampler(fits$linreg[[2]]),
                 bridge_sampler(fits$linreg[[1]]))
  if(bf$bf < ifelse(bf$log,log(3.2),3.2)){# Jeffrey's criterion
    fits$linreg = 
    fits$linreg$untransformed
    transformation = NULL
  }else{
    fits$linreg = 
    fits$linreg[[2]]
  }
  
}else{
  fits$linreg = 
    fits$linreg$untransformed
}

```

```{r}
if(is.null(transformation)){
  
  tibble(Covariate = 
         names(coef(fits$linreg)),
       Estimate = coef(fits$linreg)) %>% 
  bind_cols(as_tibble(posterior_interval(fits$linreg,
                                         prob = 0.95,
                                         pars = names(coef(fits$linreg))))) %>% 
    flextable() %>%  
    autofit() %>% 
    colformat_double(digits = 3) %>% 
    theme_booktabs()
  
  cat("The regression coefficients are reported in terms of the change in the mean of the outcome due to a unit increase in the covariate, holding all other covariates constant.")
  
}else{
  
  if(transformation == "log"){
    
    covariate_names = 
      setdiff(names(coef(fits$linreg)),
              "(Intercept)")
    tibble(Covariate = 
             covariate_names,
           Estimate = coef(fits$linreg)[-1]) %>% 
      bind_cols(as_tibble(posterior_interval(fits$linreg,
                                             prob = 0.95,
                                             pars = covariate_names ))) %>% 
      mutate(decrease_est = Estimate < 0,
             decrease_lb = `2.5%` < 0,
             decrease_ub = `97.5%` < 0) %>% 
      mutate(Estimate = 
               ifelse(decrease_est,
                      paste(round((1.0 - exp(Estimate)) * 100,1),
                            "% decrease",
                            sep=""),
                      paste(round((exp(Estimate) - 1.0) * 100,1),
                      "% increase",
                      sep="")),
             `95% CI` =
               ifelse(!decrease_lb,
                      paste("(",
                            round((exp(`2.5%`) - 1.0) * 100,1),
                            "% increase, ",
                            round((exp(`97.5%`) - 1.0) * 100,1),
                            "% increase)",
                            sep = ""),
                      ifelse(decrease_ub,
                             paste("(",
                                   round((1.0 - exp(`97.5%`)) * 100,1),
                                   "% decrease, ",
                                   round((1.0 - exp(`2.5%`)) * 100,1),
                                   "% decrease)",
                                   sep = ""),
                              paste("(",
                                    round((1.0 - exp(`2.5%`)) * 100,1),
                                    "% decrease, ",
                                    round((exp(`97.5%`) - 1.0) * 100,1),
                                    "% increase)",
                                    sep = "")))
             ) %>% 
      select(Covariate, Estimate, `95% CI`) %>% 
      flextable() %>% 
      autofit() %>% 
      colformat_double(digits = 3) %>% 
      theme_booktabs() %>% 
      print()
  
      cat("The regression coefficients are reported in terms of the % change in the mean of the outcome due to a unit increase in the covariate, holding all other covariates constant.")
  
    }else{
      tibble(Covariate = 
           names(coef(fits$linreg)),
         Estimate = coef(fits$linreg)) %>% 
    bind_cols(as_tibble(posterior_interval(fits$linreg,
                                           prob = 0.95,
                                           pars = names(coef(fits$linreg))))) %>% 
        flextable() %>%  
        autofit() %>% 
        colformat_double(digits = 3) %>% 
        theme_booktabs()
  }
}
```

## Count

### Simulate data
```{r}
sim_data$crosssec_regr$count = 
  tibble(x1 = rnorm(100),
         x2 = rnorm(100)) %>% 
  mutate(y = rnbinom(100,
                     mu = 10 * exp(0.25 * x1),
                     size = 5))
```

### Analysis - Generalized linear regression
```{r results = "hide"}
set.seed(2024)
fits$count = list()
fits$count$poisson = 
  stan_glm(y ~ x1 + x2,
           data = sim_data$crosssec_regr$count,
           family = poisson,
           diagnostic_file = file.path(tempdir(), "delete_me-1.csv"))

fits$count$nb = 
  stan_glm(y ~ x1 + x2,
           data = sim_data$crosssec_regr$count,
           family = neg_binomial_2,
           diagnostic_file = file.path(tempdir(), "delete_me-2.csv"))

bf = 
  bayes_factor(bridge_sampler(fits$count$nb),
               bridge_sampler(fits$count$poisson))

if(bf$bf < ifelse(bf$log,log(3.2),3.2)){# Jeffrey's criterion
  fits$count = 
    fits$count$poisson
}else{
  fits$count = 
    fits$count$nb
}
```

```{r}
covariate_names = 
  setdiff(names(coef(fits$count)),
          "(Intercept)")
tibble(Covariate = 
         covariate_names,
       Estimate = coef(fits$count)[-1]) %>% 
  bind_cols(as_tibble(posterior_interval(fits$count,
                                         prob = 0.95,
                                         pars = covariate_names ))) %>% 
  mutate(decrease_est = Estimate < 0,
         decrease_lb = `2.5%` < 0,
         decrease_ub = `97.5%` < 0) %>% 
  mutate(Estimate = 
           ifelse(decrease_est,
                  paste(round((1.0 - exp(Estimate)) * 100,1),
                        "% decrease",
                        sep=""),
                  paste(round((exp(Estimate) - 1.0) * 100,1),
                        "% increase",
                        sep="")),
         `95% CI` =
           ifelse(!decrease_lb,
                  paste("(",
                        round((exp(`2.5%`) - 1.0) * 100,1),
                        "% increase, ",
                        round((exp(`97.5%`) - 1.0) * 100,1),
                        "% increase)",
                        sep = ""),
                  ifelse(decrease_ub,
                         paste("(",
                               round((1.0 - exp(`97.5%`)) * 100,1),
                               "% decrease, ",
                               round((1.0 - exp(`2.5%`)) * 100,1),
                               "% decrease)",
                               sep = ""),
                         paste("(",
                               round((1.0 - exp(`2.5%`)) * 100,1),
                               "% decrease, ",
                               round((exp(`97.5%`) - 1.0) * 100,1),
                               "% increase)",
                               sep = "")))
  ) %>% 
  select(Covariate, Estimate, `95% CI`) %>% 
  flextable() %>% 
  autofit() %>% 
  colformat_double(digits = 3) %>% 
  theme_booktabs() %>% 
  print()

cat("The regression coefficients are reported in terms of the % change in the mean of the outcome due to a unit increase in the covariate, holding all other covariates constant.")
```

## Binary

### Simulate data
```{r}
sim_data$crosssec_regr$binary =
  tibble(x1 = rnorm(250),
         x2 = rnorm(250)) %>%
  mutate(y = rbinom(250,
                    1,
                    expit(-2.0 + 0.5 * x1)))
```

### Analysis - Generalized linear regression
```{r results = "hide"}
fits$binary = 
  stan_glm(y ~ x1 + x2,
           data = sim_data$crosssec_regr$binary,
           family = binomial)
```

```{r}
covariate_names = 
  setdiff(names(coef(fits$binary)),
          "(Intercept)")
tibble(Covariate = 
         covariate_names,
       Estimate = coef(fits$count)[-1]) %>% 
  bind_cols(as_tibble(posterior_interval(fits$count,
                                         prob = 0.95,
                                         pars = covariate_names ))) %>% 
  mutate_if(is.numeric,exp) %>% 
  flextable() %>% 
  autofit() %>% 
  colformat_double(digits = 3) %>% 
  theme_booktabs() 

cat("These regression results can be interpreted in terms of the odds ratio due to a unit increase in the covariate, keeping all other covariates constant.")
```


## Ordinal

### Simulate data
```{r}
sim_data$crosssec_regr$ordinal =
  tibble(x1 = rnorm(250),
         x2 = rnorm(250)) %>%
  mutate(cum_prob_cat1 = expit(-1 - 0.5 * x1),
         cum_prob_cat2 = expit(0 - 0.5 * x1),
         cum_prob_cat3 = expit(1 - 0.5 * x1),
         cum_prob_cat4 = 1.0) %>% 
  mutate(prob_cat1 = cum_prob_cat1,
         prob_cat2 = cum_prob_cat2 - cum_prob_cat1,
         prob_cat3 = cum_prob_cat3 - cum_prob_cat2,
         prob_cat4 = cum_prob_cat4 - cum_prob_cat3,
         y = 0L)
for(i in 1:nrow(sim_data$crosssec_regr$ordinal)){
  sim_data$crosssec_regr$ordinal$y[i] = 
    sample(4,1,
           prob = 
             sim_data$crosssec_regr$ordinal[i,] %>% 
             select(prob_cat1:prob_cat4) %>% 
             unlist())
}
sim_data$crosssec_regr$ordinal %<>%
  select(x1,x2,y)
```

### Analysis - Proportional odds logistic regression
```{r results = "hide"}
set.seed(2024)
fits$ordinal = 
  stan_polr(y ~ x1 + x2,
            data = 
              sim_data$crosssec_regr$ordinal %>% 
              mutate(y = factor(y,ordered = TRUE)),
            prior = R2(0.5,
                       what = "mean"))
```

```{r}
tibble(Covariate = 
         names(coef(fits$ordinal)),
       Estimate = coef(fits$ordinal)) %>% 
  bind_cols(as_tibble(posterior_interval(fits$ordinal,
                                         prob = 0.95,
                                         pars = names(coef(fits$ordinal)) ))) %>% 
  mutate_if(is.numeric,exp) %>% 
  flextable() %>% 
  autofit() %>% 
  colformat_double(digits = 3) %>% 
  theme_booktabs() 

cat("These regression results can be interpreted in terms of odds ratios corresponding to the odds of having a higher value of the outcome due to a unit increase in the covariate, keeping all other covariates constant.")
```


# Retro cohort regression

## Simulate data
```{r}
sim_data$retro_regr = 
  tibble(subject_id = 
           rep(1:100, each = 5),
         time = rep(1:5,100),
         x1 = rep(rnorm(100),each = 5),
         x2 = rep(rnorm(100),each = 5),
         re_compound_symmetry = rep(rnorm(100),each = 5)) %>% 
  group_by(subject_id) %>% 
  mutate(y = rnorm(n(),
                   mean = 2 + x1 + re_compound_symmetry - 0.1 * time^2)) %>% 
  ungroup()
```

## Analysis - Mixed effects regression
```{r results = "hide"}
set.seed(2024)
fits$retro_regr = 
  stan_glmer(y ~ time + I(time^2) + x1 + x2 + (1 | subject_id),
             data = sim_data$retro_regr)
```

```{r}
tibble(Covariate = 
         names(fixef(fits$retro_regr)),
       Estimate = fixef(fits$retro_regr)) %>% 
  bind_cols(as_tibble(posterior_interval(fits$retro_regr,
                                         prob = 0.95,
                                         pars = names(fixef(fits$retro_regr)) ))) %>% 
  flextable() %>% 
  autofit() %>% 
  colformat_double(digits = 3) %>% 
  theme_booktabs() 
```



# Prospective cohort regression

Same as above.


# Interrupted time-series

## Simulate data
```{r}
sim_data$itt = 
  tibble(time = 1:200,
         intervention = rep(c("No","Yes"),
                            c(50,150)),
         x = rnorm(200),
         y = 0.0)
for(tt in 2:nrow(sim_data$itt)){
  sim_data$itt$y[tt] = 
    1.0 * (sim_data$itt$intervention[tt] == "Yes")  + 
    0.0 * sim_data$itt$x[tt] +
    sim_data$itt$time[tt] / nrow(sim_data$itt) +
    0.5 * (sim_data$itt$y[tt - 1] -
             0.0 * sim_data$itt$x[tt] +
             sim_data$itt$time[tt] / nrow(sim_data$itt)
             ) +
    rnorm(1)
}
```

## Visualize data
```{r}
plot(sim_data$itt$y,
     type = 'l',
     bty = 'l',
     lwd = 3,
     xlab = "Time",
     ylab = "Outcome",
     cex.axis = 2,
     cex.lab = 2)
polygon(x = c(min(which(sim_data$itt$intervention == "Yes")),
              nrow(sim_data$itt),
              nrow(sim_data$itt),
              min(which(sim_data$itt$intervention == "Yes"))),
        y = c(min(sim_data$itt$y),
              min(sim_data$itt$y),
              max(sim_data$itt$y),
              max(sim_data$itt$y)),
        border = NA,
        col = adjustcolor("steelblue",0.5))
axis(3,
     at = median(which(sim_data$itt$intervention == "No")),
     labels = "Pre-intervention",
     cex.axis = 2,
     tick = FALSE,
     col.axis = gray(0.4))
axis(3,
     at = median(which(sim_data$itt$intervention == "Yes")),
     labels = "Post-intervention",
     cex.axis = 2,
     tick = FALSE,
     col.axis = "steelblue")
```

## Analysis - Time series
```{r results = "hide"}
set.seed(2024)
X_matrix = 
  model.matrix(~intervention + x + time,
               data =
                 sim_data$itt %>%
                 mutate_if(is.numeric,scale))[,-1]
fits$itt = list()
fits$itt$ar1 = 
  stan_sarima(drop(scale(sim_data$itt$y)),
              order = c(1,0,0),
              xreg = X_matrix,
              prior_breg = bayesforecast::normal(0,2.5))
fits$itt$ma1 =
  stan_sarima(drop(scale(sim_data$itt$y)),
              order = c(0,1,0),
              xreg = X_matrix,
              prior_breg = bayesforecast::normal(0,2.5))
  
bf = 
  bayes_factor(fits$itt$ar1,
               fits$itt$ma1)

if(bf$bf > ifelse(bf$log,log(1),1)){# Jeffrey's criterion
  fits$itt = 
    fits$itt$ar1
}else{
  fits$itt = 
    fits$itt$ma1
}

```

```{r}
summary(fits$itt,
        prob = 0.95)[grep("breg",
                       row.names(summary(fits$itt))),
                  c("mean","2.5%","97.5%")] %>% 
  as_tibble() %>% 
  mutate_all(~ .x * 
               sd(sim_data$itt$y) / 
               apply(model.matrix(~intervention + x + time,
                                  data = sim_data$itt)[,-1],
                     2,sd)) %>%
  rename(Estimate = mean) %>% 
  mutate(Covariate = colnames(X_matrix)) %>% 
  relocate(Covariate)
```



# Stepped-wedge

## Simulate data
```{r}
sim_data$wedge = 
  tibble(subject_id = 
           rep(1:100, each = 6),
         time = rep(1:6,100),
         x1 = rep(rnorm(100),each = 6),
         x2 = rep(rnorm(100),each = 6),
         re_compound_symmetry = rep(rnorm(100),each = 6),
         treatment = 
           c(rep(rep(0:1,c(1,5)), 20),
             rep(rep(0:1,c(2,4)), 20),
             rep(rep(0:1,c(3,3)), 20),
             rep(rep(0:1,c(4,2)), 20),
             rep(rep(0:1,c(5,1)), 20))) %>% 
  mutate(y = treatment + 0.5 * x1 - time + re_compound_symmetry + rnorm(n()))

# Check to make sure data was produced correctly:
sim_data$wedge %>% 
  group_by(time) %>% 
  summarize(s = sum(treatment))
```

## Analysis - Mixed effects regression
```{r results = "hide"}
set.seed(2024)
fits$wedge = 
  stan_glmer(y ~ treatment + x1 + x2 + time + (1|subject_id),
             data = sim_data$wedge)
```

```{r}
tibble(Covariate = 
         names(fixef(fits$wedge)),
       Estimate = fixef(fits$wedge)) %>% 
  bind_cols(as_tibble(posterior_interval(fits$wedge,
                                         prob = 0.95,
                                         pars = names(fixef(fits$wedge)) ))) %>% 
  flextable() %>% 
  autofit() %>% 
  colformat_double(digits = 3) %>% 
  theme_booktabs() 
```



# Cluster randomized

## Simulate data
```{r}
sim_data$CRT = 
  tibble(cluster_id = 
           rep(1:20, each = 10),
         x1 = rnorm(200),
         x2 = rnorm(200),
         re_compound_symmetry = rep(rnorm(20),each = 10),
         treatment = rep(sample(rep(0:1,10)),each = 10)) %>% 
  mutate(y = treatment + 0.5 * x1 + re_compound_symmetry + rnorm(n()))

# Check to make sure data was produced correctly:
sim_data$CRT %$% 
  table(cluster_id,treatment)
```

## Analysis - Mixed effects regression
```{r results = "hide"}
set.seed(2024)
fits$CRT = 
  stan_glmer(y ~ x1 + x2 + treatment + (1|cluster_id),
             data = sim_data$CRT)
```

```{r}
tibble(Covariate = 
         names(fixef(fits$CRT)),
       Estimate = fixef(fits$CRT)) %>% 
  bind_cols(as_tibble(posterior_interval(fits$CRT,
                                         prob = 0.95,
                                         pars = names(fixef(fits$CRT)) ))) %>% 
  flextable() %>% 
  autofit() %>% 
  colformat_double(digits = 3) %>% 
  theme_booktabs() 
```


# RCT

## Simulate data
```{r}
sim_data$RCT = 
  tibble(treatment = rep(0:1,each = 50)) %>% 
  mutate(y = treatment + rnorm(100))
```

## Analysis - ANOVA
You can also adjust for covariates, but not necessary.
```{r results = "hide"}
set.seed(2024)
fits$RCT = 
  stan_glm(y ~ treatment,
           data = sim_data$RCT)
```

```{r}
tibble(Covariate = 
         names(coef(fits$RCT)),
       Estimate = coef(fits$RCT)) %>% 
  bind_cols(as_tibble(posterior_interval(fits$RCT,
                                         prob = 0.95,
                                         par = names(coef(fits$RCT))))) %>% 
  flextable() %>% 
  colformat_double(digits = 3) %>% 
  theme_booktabs()
```


# One last tip
It may be useful to you to create your code through string manipulation and then evaluate it.  I can redo the previous code for the CRT as follows:
```{r results = "hide"}
response_variable = "y"
treatment_variable = "treatment"
covariates = c("x1","x2")
cluster_variable = "cluster_id"
data_name = "sim_data$CRT"
set.seed(2024)
fits$CRT = 
  eval(parse(text = 
               paste0(
                "stan_glmer(",
                response_variable,
                "~",
                 paste(c(treatment_variable,
                         covariates),
                       collapse = " + "),
                " + (1|",
                cluster_variable,
                "), data = ",
                data_name,
                ")"
               )))

```

```{r}
tibble(Covariate = 
         names(fixef(fits$CRT)),
       Estimate = fixef(fits$CRT)) %>% 
  bind_cols(as_tibble(posterior_interval(fits$CRT,
                                         prob = 0.95,
                                         pars = names(fixef(fits$CRT)) ))) %>% 
  flextable() %>% 
  autofit() %>% 
  colformat_double(digits = 3) %>% 
  theme_booktabs() 
```




