---
title: "BPVal"
author: "AN"
date: "2026-06-17"
output: html_document
---

Mixed effects models can be used to incorporate grouping variables in
the context of generalized linear models. These types of models
incorporate both fixed and random effects on the mean of the response distribution
through the link function. That is,

$g(\mu_i) = \mathbf{x}_i^\top \mathbf{\beta} + \mathbf{z}_i^\top \mathbf{u}$

$E[y_i | \mathbf{u}] = \mu_i$

where $y_i |\mathbf{u}$ is distributed according to an exponential family. 
We are interested in the chisquare-type pearson statistic: 

$T(\mathbf{y},\theta) = \frac{\sum_i(y_i - E[y_i|\theta])^2}{V[y_i|\theta]}$


The `rstanarm` package is capable of fitting generalized linear mixed
effects models, from which we can obtain samples of the linear predictor
(incorporating both the fixed and random effect) from the posterior distribution.

After fitting an `rstanarm` model, we have access to the posterior distribution
of conditional expectations `posterior_epred()`, the posterior distribution of linear predictors `posterior_linpred()`, and the
posterior predictive distribution `posterior_predict()`. Note that
in the GLM framework, the conditional expectations are related to the posterior linear predictors,
and so the conditional expectations are accessed directly without having to transform the linear
predictor through the link function. Denoting a posterior draw of theta as $\theta^{(d)}$, our statistic
of interest is:

\begin{align*}
  T^{(d)}(\mathbf{y}, \mathbf{\beta}) &= \sum_{i=1}^n \frac{(y_i - E[y_i|\theta^{(d)}])^2}{V[y_i|\theta^{(d)}]} \\
\end{align*}




The variance of $y_i$ conditioned on $\theta$ is dependent on the distribution
of $y|\theta$. The following are these variance quantities:

For Gaussian:

\begin{align*}
  E[y_i|\theta^{(d)}] &= \mu_i^{(d)} \\
  V[y_i|\theta^{(d)}] &= \sigma_i^{2{(d)}} \\
\end{align*}

For Binomial:

\begin{align*}
  E[y_i|\theta^{(d)}] &= t_i p_i^{(d)} \\
  V[y_i|\theta^{(d)}] &= t_i p_i^{(d)} (1 - p_i^{(d)}) \\
\end{align*}

For Poisson:

\begin{align*}
  E[y_i|\theta^{(d)}] &= \lambda_i^{(d)} \\
  V[y_i|\theta^{(d)}] &= \lambda_i^{(d)} \\
\end{align*}

```{r, echo = FALSE}

library(lubridate)
library(tidyverse)
library(rstanarm)
library(lattice)
data(sleepstudy, package = "lme4")
data(cbpp, package = "lme4")


```

# Preemptive Check
Checking that we can use posterior_epred() instead of transforming posterior_linpred()

```{r, include = FALSE}

# Confirming that conditional expectations are equivalent to applying inverse link
# on linear predictors
fit =
  stan_glmer(cbind(incidence) ~ period + (1 | herd),
             data = cbpp, family = poisson)



# Create conditional expectations in 3 ways:

E_1 <- invlogit(posterior_linpred(fit)) # "manually" transforming eta

E_2 <- posterior_linpred(fit, transform = TRUE) # Using built in transformation of linpred

E_3 <- posterior_epred(fit) # Using built in conditional expectation


# Check equivalence

mean(E_1 == E_2) # Sometimes equal?

mean(E_2 == E_3) # Equal!


# Check if this is due to rounding error

mean(abs(E_1 - E_2) < 0.00000000001) # Always within a small amount, package implementation is probably more sophisticated

```

# Function

```{r, include = FALSE}


pearson_bpval <- function(fit, family, draws){
  if (family == "binomial"){
    
    # Calculate number of trials
    trials <- rowSums(fit$y)
    
    # Set up matrices of d x n for observed and simulated data
    y_obs <- matrix(rep(fit$y[,1], draws), ncol = length(fit$y[,1]), nrow = draws, byrow = TRUE) 
    y_new <- posterior_predict(fit, draws = draws)
    
    # Calculate conditional probabilty
    p <- posterior_epred(fit, draws = draws)
    
    # Calculate expectation
    E <- trials * p
    
    # Calculate variance
    V <- trials * p * (1-p)
    
    # Calculate vectors of the statistic
    T_obs <- rowSums((y_obs - E)^2 / V)
    T_new <- rowSums((y_new - E)^2 / V)
    
    pval <- mean(T_obs > T_new)
    
    plot(T_obs, T_new)
    abline(a = 0, b = 1)

    
    return(pval)
    
  }
  
    if (family == "gaussian"){
    
    # Set up matrices of d x n for observed and simulated data
    y_obs <- matrix(rep(fit$y, draws), ncol = length(fit$y), nrow = draws, byrow = TRUE) 
    y_new <- posterior_predict(fit, draws = draws)
    
    # Calculate conditional expectation
    E <- posterior_epred(fit, draws = draws)
    
    # Calculate conditional variance
    V <- as.matrix(fit)[, "sigma"][1:draws]^2
    
    # Calculate vectors of the statistic
    T_obs <- rowSums((y_obs - E)^2 / V)
    T_new <- rowSums((y_new - E)^2 / V)
    
    pval <- mean(T_obs > T_new)
    
    plot(T_obs, T_new)
    abline(a = 0, b = 1)
    
    return(pval)
    
    }
    
    if (family == "poisson"){
    
    # Set up matrices of d x n for observed and simulated data
    y_obs <- matrix(rep(fit$y, draws), ncol = length(fit$y), nrow = draws, byrow = TRUE) 
    y_new <- posterior_predict(fit, draws = draws)
    
    # Calculate conditional expectation
    E <- posterior_epred(fit, draws = draws)
    
    # Calculate conditional variance
    V <- posterior_epred(fit, draws = draws)
    
    # Calculate vectors of the statistic
    T_obs <- rowSums((y_obs - E)^2 / V)
    T_new <- rowSums((y_new - E)^2 / V)
    
    pval <- mean(T_obs > T_new)
    
    plot(T_obs, T_new)
    abline(a = 0, b = 1)

    
    return(pval)
    
    }
  
}




```


# Testing
## "Good" Example

```{r, include = FALSE}

library(rstanarm)
set.seed(123)
data(sleepstudy, package = "lme4")

# ------------------------------------------------------------------------------

# Testing on a real dataset
fit =
  stan_glmer(Reaction ~ Days + (Days | Subject),
             data = sleepstudy, family = gaussian)


draws = 4000

# ------------------------------------------------------------------------------

# Set up matrices of d x n for observed and simulated data
  y_obs <- matrix(rep(fit$y, draws), ncol = length(fit$y), nrow = draws, byrow = TRUE) 
  y_new <- posterior_predict(fit, draws = draws)
    
# Calculate conditional expectation
  E <- posterior_epred(fit, draws = draws)
    
# Calculate conditional variance
  V <- as.matrix(fit)[, "sigma"][1:draws]^2
    
# Calculate vectors of the statistic
  T_obs <- rowSums((y_obs - E)^2 / V)
  T_new <- rowSums((y_new - E)^2 / V)
    
    pval <- mean(T_obs > T_new)
    
    plot(T_obs, T_new)
    abline(a = 0, b = 1)
    
    pval <- mean(T_obs > T_new)
    print(pval) # Around 0.5, expected



```



## "Bad" Example
```{r, include = FALSE}
# Succinct example
library(rstanarm)
set.seed(123)



# ------------------------------------------------------------------------------
# 10 groups of 10 each, outcome is completely independent of x and group structure
mydat_bad <- data.frame(
  y = runif(n = 100, min = 0, max = 10),
  group = factor(rep(1:10, each = 10)),
  x = runif(n = 100, min = -5, max = 5))

# The outcome is uniform, not normal
plot(mydat_bad$y)

# Create fit
myfit_bad =
  stan_glmer(y ~ x + (1 | group),
             data = mydat_bad, family = "gaussian")

# ------------------------------------------------------------------------------

draws = 4000
fit <- myfit_bad

# Set up matrices of d x n for observed and simulated data
  y_obs <- matrix(rep(fit$y, draws), ncol = length(fit$y), nrow = draws, byrow = TRUE) 
  y_new <- posterior_predict(fit, draws = draws)
    
# Calculate conditional expectation
  E <- posterior_epred(fit, draws = draws)
    
# Calculate conditional variance
  V <- as.matrix(fit)[, "sigma"][1:draws]^2
    
# Calculate vectors of the statistic
  T_obs <- rowSums((y_obs - E)^2 / V)
  T_new <- rowSums((y_new - E)^2 / V)
    
    pval <- mean(T_obs > T_new)
    
    plot(T_obs, T_new)
    abline(a = 0, b = 1)
    
    pval <- mean(T_obs > T_new)
    print(pval) # Around 0.5, despite outcome indep of x and not even normal



```






# Mean Structure Focus

Here, we assess model fit using Bayes P-values based on loess curves
fit to the predicted residuals and observed residuals.

Define $y_i^{(d)}$ as the $i$th posterior predicted value from the $d$th draw of $\theta$. Similarly, define $\mu_i^{(d)}$ as the  




```{r, include = FALSE}

library(lubridate)
library(tidyverse)
library(rstanarm)
library(patchwork)

N = 200

set.seed(2026)
data = 
  list(
    good = 
      tibble(x = runif(N)) |>
      mutate(gr = rep(1:10,each = n() / 10),
             y = x + rnorm(n())),
    
    quad_normal = 
      tibble(x = runif(N)) |>
      mutate(gr = rep(1:10,each = n() / 10),
             y = x + x^2 + rnorm(n())),
    
    very_quad_normal = 
      tibble(x = runif(N)) |>
      mutate(gr = rep(1:10,each = n() / 10),
             y = x + 25*x^2 + rnorm(n())),
    
    quad_unif = 
      tibble(x = runif(N)) |>
      mutate(gr = rep(1:10,each = n() / 10),
             y = x + 3*x^2 + runif(n())),
    
    none_normal = 
      tibble(x = runif(N)) |>
      mutate(gr = rep(1:10,each = n() / 10),
             y = rnorm(n())),
    
    none_unif = 
      tibble(x = runif(N)) |>
      mutate(gr = rep(1:10,each = n() / 10),
             y = runif(n()))
  )

vis <- lapply(
  names(data),
  function(j){
    data[[j]] |> 
      ggplot() +
      geom_point(aes(y=y,x=x)) +
      theme_minimal(base_size = 18) +
      ggtitle(j)
  }
) |> 
  wrap_plots()
  

fits =
  lapply(
    data,
    function(X){
      stan_lmer(y ~ x + (1 | gr),
                data = X,
                cores = 4)
    })
beepr::beep(3)


lmer_bpv = function(fit){
  # Get response variable
  y = 
    fit$data[,all.vars(fit$formula)[1]]
  
  # Get draws of fitted values
  E_y = 
    posterior_epred(fit)
  
  # Get draws of residuals
  res = 
    matrix(y,
           nrow(E_y),
           NROW(y),
           byrow = T) - E_y
  
  # Get draws of response
  y_pred = 
    posterior_predict(fit)
  
  # Get residuals for response draws
  res_pred = 
    y_pred - E_y
  
  
  # Create helper function to fit loess
  helper = function(i){
     l1 = 
       loess(res[i,] ~ E_y[i,],
             span = 0.6) |> 
       predict()
     l2 = 
       loess(res_pred[i,] ~ E_y[i,],
             span = 0.6) |> 
       predict()
     
     sum(l1^2) > sum(l2^2)
  }
  
  # Compute bayesian p-value
  mean(sapply(1:nrow(res),helper))
}

bpv =
  lapply(fits,
         lmer_bpv)
beepr::beep(3)
bpv



x <- runif(n = 10, 0, 10)
beta <- 2
y <- x * beta + rnorm(n = 10, sd = 10)

mymod <- loess(y ~ x)

plot(x, predict(mymod), type = "l")




```
