---
title: "lme_bpv Demo"
format: html
author: AN
---




# Mean Structure Focus

Here, we assess model fit using Bayes P-values based on loess curves
fit to the predicted residuals and observed residuals.

Define $y_i^{(d)}$ as the $i$th posterior predicted value from the $d$th posterior draw of $\theta$. Similarly, define $\mu_i^{(d)} = E[y_i|\theta^{(d)}]$ as the expected value of $y_i$ under the $d$th posterior draw of $\theta$.

Consider the $n$ observed residuals and the $n$ predicted residuals:

\begin{align*}
  r_{i, obs}^{(d)} = y_i - \mu_i^{(d)} \\
  r_{i, pred}^{(d)} = y_i^{(d)} - \mu_i^{(d)} \\
\end{align*}

If we fit two loess curves:

\begin{align*}

  \text{LOESS}(\mathbf{r}_{obs}^{(d)}, \boldsymbol{\mu}^{(d)}) \\
  \text{LOESS}(\mathbf{r}_{pred}^{(d)}, \boldsymbol{\mu}^{(d)}) \\

\end{align*}

These two models give rise to the predicted residuals: $\hat{\mathbf{r}}_{obs}^{(d)}$
and $\hat{\mathbf{r}}_{pred}^{(d)}$. Define the statistics

\begin{align*}
  T_{obs}^{(d)} = \frac{1}{n}\sum_{i = 1}^n \left(\hat{r}_{obs}^{(d)}\right)^2 \\ 
  T_{pred}^{(d)} = \frac{1}{n}\sum_{i = 1}^n \left(\hat{r}_{pred}^{(d)}\right)^2 \\
\end{align*}

The bayesian p-value is given by

\begin{align*}
  \frac{1}{D}\sum_{d = 1}^{D} I(T_{pred}^{(d)} > T_{obs}^{(d)})
\end{align*}

# Function

```{r}


lme_bpv <- function(object){

  # Response
  y_obs <- object$data[,all.vars(object$formula)[1]]

  # Predicted expectations
  mu <- rstanarm::posterior_epred(object)

  # Response vs fitted residuals
  res_obs <- matrix(data = y_obs,
                nrow = nrow(mu),
                ncol = NROW(y_obs),
                byrow = T) - mu

  # Predicted response
  y_pred <- posterior_predict(object)

  # Predicted response vs fitted residuals
  res_pred <- y_pred - mu

  loess_obs <- function(i){

    # Fit loess to residuals vs fitted
      loess_fit = loess(res_obs[i,] ~ mu[i,]) |> predict()

      sum(loess_fit^2)
      
  }

  loess_pred <- function(i){

    # Fit loess to residuals vs fitted
    loess_fit = loess(res_pred[i,] ~ mu[i,]) |> predict()

    sum(loess_fit^2)

  }

  T_obs <- sapply(1:nrow(res_obs), loess_obs)
  T_pred <- sapply(1:nrow(res_pred), loess_pred)

  
  rslt <- list(
    bpval = mean(T_obs > T_pred),
    bpplot = 
      ggplot(data.frame(T_obs, T_pred)) + 
      geom_point(mapping = aes(x = T_obs, y = T_pred)) +
      geom_abline(slope = 1, intercept = 0)
  )
  
  return(rslt)

  }





```

# Example

```{r}

library(tidyverse)
library(rstanarm)

# Model fit

N = 200

good =
  tibble(x = runif(N)) |>
  mutate(gr = rep(1:10,each = n() / 10),
         y = x + rnorm(n()))

gfit <- stan_lmer(y ~ x + (1 | gr), data = good)

# ------------------------------------------------------------------------------

  # Response
  y_obs <- gfit$data[,all.vars(gfit$formula)[1]]

  # Predicted expectations
  mu <- rstanarm::posterior_epred(gfit)

  # Response vs fitted residuals
  res_obs <- matrix(data = y_obs,
                nrow = nrow(mu),
                ncol = NROW(y_obs),
                byrow = T) - mu

  # Predicted response
  y_pred <- posterior_predict(gfit)

  # Predicted response vs fitted residuals
  res_pred <- y_pred - mu

  # ----------------------------------------
  # LOESS curves
  
    # Fit loess to residuals vs fitted
      loess_fit_1 = loess(res_obs[1,] ~ mu[1,]) |> predict()
      sum(loess_fit_1^2)

    # Fit loess to residuals vs fitted
      loess_fit_2 = loess(res_pred[1,] ~ mu[1,]) |> predict()
      sum(loess_fit_2^2)

    # Show loess curves
    plot(mu[1,], res_obs[1,], col = "blue", ylab = "Residuals")
    points(mu[1,], res_pred[1,] , col = "red")
    data.frame(x = mu[1,], y = loess_fit_1) |> arrange(x) |> lines(, col = "blue")
    data.frame(x = mu[1,], y = loess_fit_2) |> arrange(x) |> lines(, col = "red")
    legend("topleft",legend = c("obs","pred"), col = c("blue","red"), fill = c("blue","red"))
    
    # Isolate curves
    
    plot(mu[1,], loess_fit_1, col = "blue", ylab = "Residuals")
    points(mu[1,], loess_fit_2, col = "red")

    # Square residuals
    plot(mu[1,], loess_fit_1^2, col = "blue", ylab = "Squared Residuals")
    points(mu[1,], loess_fit_2^2, col = "red")
    
    lme_bpv(gfit)
  

  











```


# Dan's Code

```{r}


bpv_d = function(fit){
  # Get response variable
  y = 
    fit$data[,all.vars(fit$formula)[1]]
  
  # Get draws of fitted values
  E_y = 
    posterior_epred(fit)
  
  # Get draws of residuals
  res = 
    matrix(y,
           nrow(E_y),
           NROW(y),
           byrow = T) - E_y
  
  # Get draws of response
  y_pred = 
    posterior_predict(fit)
  
  # Get residuals for response draws
  res_pred = 
    y_pred - E_y
  
  
  # Create helper function to fit loess
  helper = function(i){
     l1 = 
       loess(res[i,] ~ E_y[i,],
             span = 0.6) |> 
       predict()
     l2 = 
       loess(res_pred[i,] ~ E_y[i,],
             span = 0.6) |> 
       predict()
     
     sum(l1^2) > sum(l2^2)
  }
  
  # Compute bayesian p-value
  mean(sapply(1:nrow(res),helper))
}

```

# Testing/Simulation

Testing bpv "rejection rate", (if term is applicable). That is,
determine how often under a linear mixed model, the model is deemed
"not fitting", under correct/incorrect model specifications.


```{r}

lin_mod <- vector()
none_unif_mod <- vector()
very_quad_normal_mod <- vector()

lin =
  tibble(x = runif(N)) |>
  mutate(gr = rep(1:10,each = n() / 10),
         y = x + rnorm(n()))

none_unif = 
      tibble(x = runif(N)) |>
      mutate(gr = rep(1:10,each = n() / 10),
             y = runif(n()))

very_quad_normal = 
      tibble(x = runif(N)) |>
      mutate(gr = rep(1:10,each = n() / 10),
             y = x + 25*x^2 + rnorm(n()))

for (i in 1:80){
  
N = 200

lin =
  tibble(x = runif(N)) |>
  mutate(gr = rep(1:10,each = n() / 10),
         y = x + rnorm(n()))

fit1 <- stan_lmer(y ~ x + (1 | gr), data = good, refresh = 0)
fit2 <- stan_lmer(y ~ x + (1 | gr), data = none_unif, refresh = 0)
fit3 <- stan_lmer(y ~ x + (1 | gr), data = very_quad_normal, refresh = 0)

lin_mod[i] <- lme_bpv(fit1)$bpval
none_unif_mod[i] <- lme_bpv(fit2)$bpval
very_quad_normal_mod[i] <- lme_bpv(fit3)$bpval


}


plot(lin_mod)
plot(none_unif_mod)
plot(very_quad_normal_mod)









```

